The Correspondence Analysis of Two Mode Networks Revisited

### 2025-02-04 main-R1.tex

\section{CA and Dual Centrality in Two-Mode Networks} \label{sec:ca}
The reader may ask what the point of going through all these reflective scoring, since \citet{bonacich1991simultaneous} already developed a dual conception of centrality in two-mode networks based on a very similar idea: Define the centralities of entities in one mode based on the centralities of entities in the other mode to which they are connected. Recall that, as noted earlier in that paper, Bonacich also considered CA in passing but dismissed it as a tool for assigning scores to nodes in two-mode networks, noting that it was better at capturing the clique structure of women who attended different events \citeyearpar[164]{bonacich1991simultaneous}. While using CA to capture patterns of similarity in attendance between persons and groups makes sense (although the exact patterns of similarity captured by CA are not reducible to ``cliques'' as traditionally understood), CA can also be used to assign meaningful scores to nodes in a social network. In fact, the first CA dimension captures the limit $(q \rightarrow \infty)$ of the reflective centralities discussed in the previous section \citep{mealy2019interpreting}.

To see this, recall that the solutions to the following system of linear equations give the Bonacich $(C^B)$ two-mode centralities:

@@

Where $A$ is the original affiliation matrix. Equations~\ref{eq:bon1} and~\ref{eq:bon2} have the typical form of a linear eigensystem, which means the unknown $C^B_g$ and $C^B_p$ scores can be obtained from the row and column eigenvectors corresponding to the largest eigenvalue $\lambda$, after performing a singular value decomposition (SVD) of the rectangular affiliation matrix $A$. Importantly, as Faust \citeyearpar[170]{faust1997centrality} notes, it is possible to express Equations~\ref{eq:bon1} and~\ref{eq:bon2} in terms of person and group-specific centralities.

@@

Where $\lambda$ is the (first) eigenvalue corresponding to the eigenvector containing the $C^B_p$ scores. These equations make clear that in the Bonacich dual-scoring approach, the centrality of people is equal to the sum of the centralities scores of the groups they join, and the centrality of groups is equivalent to the sum of the centralities of their members. This approach thus nicely captures the duality property since the centralities of nodes in each mode are given by aggregating their connections to nodes in the other mode. Note the formal similarity between equations~\ref{eq:faust1} and~\ref{eq:faust2} and equations~\ref{eq:Rq_p} and~\ref{eq:Rq_g}. The main difference is that in the reflective equations, a person's centrality is proportional to the \textit{activity-weighted} centralities of the groups they join, and a group's centrality is proportional to the \textit{group size weighted} centrality of the people who are members. We will return to this crucial point later.

\subsection{The Bonacich Two-Mode Centrality as a Reflective Centrality} \label{subsec:bonref}
To see the connection between the reflective and Bonacich two-mode centralities more clearly, we can motivate the Bonacich two-mode centrality using the same ``reflective'' approach we used to introduce the HH reflective centrality in Section~\ref{sec:ref2mode}. Admittedly, this is a somewhat unorthodox way of presenting the eigenvector centrality for two-mode networks~\citep{bonacich1991simultaneous}, but it will help clarify the similarities and differences between the HH and the Bonacich approaches. Accordingly, starting with Equations~\ref{eq:R1_p} and~\ref{eq:R1_g}, we can define a second-order ``Bonacich-reflection'' on both the persons and groups using the formulas:

@@

Where the denominator in~\ref{eq:BRq_pn} and~\ref{eq:BRq_gn} is the Euclidean vector norm.\footnote{For any vector $\mathbf{x}$ of length $n$, the $L_2$ norm is given by: $||\mathbf{x}||_2 = \sqrt{\sum_i^n x_i^2}$.} The normalization will prevent divergence of the sum of centralities for persons and groups, formalizing the weaker assumption of \textit{proportionality} between the centrality of each set of nodes and the sum of the centralities of the nodes in the other mode to which they are connected rather than the stronger assumption of strict equivalence \citep{bonacich_lloyd01}. Furthermore, the normalization guarantees convergence and the ``freezing'' of the sums of sums around steady values after a few iterations. These values will be equivalent (up to rounding error) to the (absolute value) of the dominant row and column eigenvectors of $\mathbf{A}$ as given in~\ref{eq:bon1} and~\ref{eq:bon2}.\footnote{Note that dividing by any vector norm---e.g., the $L_1$ or max norm---will prevent divergence and return scores perfectly correlated to the Bonacich eigenvector approach. Dividing by the $L_2$ norm returns scores that are \textit{exactly} the same, save for rounding error, as the absolute value of the dominant left and right eigenvectors of the affiliation matrix.} In fact, the iterative (normalized sum of sums) approach is one way of computing the leading eigenvectors of a rectangular matrix (e.g., the ``power'' method of \citet{mises1929praktische}).

This exercise establishes that there is more than a superficial similarity between the HH ``reflective'' centralities and the Bonacich two-mode centralities. In fact, both can be seen as instantiating an underlying reflective model of how centrality is distributed in the two-mode network, with the Bonacich approach presuming that centrality points are distributed equally across nodes in each mode regardless of their own centrality (both large and small degree nodes distribute the same ``amount'' of centrality to others) and the HH reflective approach normalizing by the centrality of nodes in each mode, so that nodes distribute a given centrality quantum that is proportional to their first-order degree centrality, with large degree nodes having less centrality to distribute than low degree nodes.\footnote{Note that this is precisely the difference between the Bonacich centrality in one-mode networks \citep{bonacich1972factoring} and the PageRank centrality \citep{brin1998anatomy}.}

\subsection{HH Reflections and Correspondence Analysis} \label{subsec:refeigen}
Since iterating through normalized sums of sums is one way of obtaining the Bonacich two-mode centralities, it would be surprising if the equilibrium values of the HH reflective iterations were not themselves the solution to some eigenvalue decomposition problem. As has been noted recently by \citet{mealy2019interpreting} and \citet{van2021correspondence}, the Hidalgo and Hausmann's \citeyearpar{hidalgo2009building} reflective centralities can indeed be obtained directly (without iterations) as the solution to an eigenvalue decomposition problem.

@@

\begin{equation}
    \left(AA^T\right)C^B_p = \lambda C^B_p
    \label{eq:bon3}
\end{equation}

\begin{equation}
    \left(A^TA\right)C^B_g = \lambda C^B_g
    \label{eq:bon4}
\end{equation}

This shows that the dual centrality Bonacich scores for persons and groups are equivalent to the eigenvectors of the respective one-mode projection matrices corresponding to the first (largest) eigenvalue.

Typically, when CA is presented as the solution to an eigenvalue problem, this is done by showing how the relevant scores can be obtained via the singular value decomposition (SVD) of the rectangular affiliation matrix~\cite{bonacich1991simultaneous, faust2005using}. Here, we will show that CA can also be motivated, and scores obtained via a similar ``dual'' projection just like the Bonacich eigenvector scores, except that these are scores associated with a \textit{degree-weighted} projection of the original affiliation matrix.

To see this, consider the $|P| \times |P|$ matrix $Dp$, containing the ``first order'' reflective centralities of each person $C^R(1)_p$ (activity) along the diagonals and zeroes in every other cell. In the same way, consider the $|G| \times |G|$ matrix $Dg$, which contains the ``first order'' (degree) centralities of each group $C^R(1)_g$ (size) along the diagonals and zeroes in every other cell. Using these matrices, we can compute \textit{row stochastic} versions of the affiliation matrix and its transpose (recall that a matrix is row stochastic if its rows sum to one).

@@

We can do the same with the groups, except that we pre-multiply the \textit{transpose} of the original affiliation matrix by $\mathbf{D}g^{-1}$ (which now contains the \textit{inverse} of the first-order centralities of each group $C^R_g(1)$ along the diagonals) yielding the matrix $\mathbf{P}_{gp}$ of dimensions $|G| \times |P|$:

@@

Both $\mathbf{P}_{pp}$ and $\mathbf{P}_{gg}$ are square ($|P| \times |P|$ and $|G| \times |G|$) respectively) degree-normalized projection matrices linking people and groups. Like their constituent matrices, both $\mathbf{P}_{pp}$ and $\mathbf{P}_{gg}$ are row-stochastic (rows constrained to sum to 1.0), but not symmetric, as can be seen in Figure~\ref{fig:dwproj}. Interestingly, one way to interpret the entries in the degree-weighted projection matrices for people and groups is by giving the \textit{probabilities} that a random walker starting at the row node and, following any sequence of $person-group-person'-group'$ hops, will reach the column person. Thus, higher values indicate an affinity or proximity between the people based on indirect connections in the two-mode network (and the same for groups in the corresponding matrix). The reason for the asymmetry is that depending on their pattern of connections to other groups, person $i$ may have more ways of indirectly reaching person $j$ than vice versa (unless $i$ and $j$ are connected to the same groups).

@@

With the $C^R$ scores obtained from the eigenvectors corresponding to the \textit{second} largest eigenvalue of the degree-weighted projection matrices. This is because, as we have seen, both equations~\ref{eq:ca3} and~\ref{eq:ca4} produce row stochastic matrices; this means that their largest (dominant) eigenvalue will also be 1.0, which will be associated with a trivial first eigenvector containing constant values for both the people and the groups.  Accordingly, the scores we seek, which are the same as the CA scores and the limit HH reflective centrality scores, are those stored in the eigenvectors associated with the second largest eigenvalue \citep{van2021correspondence}; these will be equivalent to the row and column scores corresponding to the first non-trivial dimension (for people and groups respectively) obtained from a simple CA of the original affiliation matrix \citep[398, eq. 9.17]{fouss2016algorithms}.\footnote{Note also that just like the eigenvector decomposition, the matrices implied by equations~\ref{eq:dam1} and~\ref{eq:dam2} have equivalent sets of (non-zero) eigenvalues equal to the rank of the original affiliation matrix.}

More importantly, because $\mathbf{P}_{pp} = Dp^{-1}ADg^{-1}A^T$ and $\mathbf{P}_{gg} = Dg^{-1}A^TDp^{-1}A$ there is a formal equivalence in the duality relations implied by equations ~\ref{eq:dam1} and~\ref{eq:dam2} and that implied by the Bonacich two-mode centralities in equations~\ref{eq:bon3} and~\ref{eq:bon4}. Both extract individual and group centralities as eigenvectors of the one-mode projection of the original affiliation matrix: $AA^T$ and $\mathbf{P}_{pp}$ for people and $A^TA$ and $\mathbf{P}_{gg}$ for groups. The key difference is that the reflective centralities pre-multiply the affiliation matrix and its transpose by the inverse of the first-order centralities of the nodes in each mode before computing the eigenvalue decomposition, essentially normalizing the one-mode projection by the degrees of both sets of nodes.

This can be clearly seen if we express $\left(Dg^{-1}\mathbf{A}^TDp^{-1}\mathbf{A}\right)$ in~\ref{eq:dam1} in terms of each cell entry \cite[eq. 4]{mealy2019interpreting}:

@@

The same reasoning applies to groups in~\ref{eq:dam2}, whose entries are given by:

@@

Now, as has been noted by two-mode data analysts in other contexts \citep[e.g.,][]{faust2005using} it turns out that this ``double-pre-weighting'' of each cell entry by the degrees of each mode (e.g., the row and column sums of the original affiliation matrix) is precisely that used for the CA of a two-mode matrix.\footnote{More accurately, cells are weighted by the inverse of the square root of the product of the row, and column sums \citep[e.g.,][124]{faust2005using}.}.\footnote{As \citet[126]{faust2005using} notes, the first non-trivial CA dimension is also given by the eigenvector corresponding to the second-largest eigenvalue of $D_p^{-1/2}\mathbf{A}D_g^{-1/2}$.} In fact, the method of ``reflections'' is just a re-discovery of the older idea of reciprocal averaging \citep{hill1973reciprocal}, which is yet another way of computing row and column scores for elements in a two-mode matrix that are substantively equivalent to those obtained by (the first non-trivial dimension of) CA \citep{mealy2019interpreting}. As we can see in Figure~\ref{fig:ca-ref-corr}, the scores computed via the iterative method of reflections ($q = 26$) correspond exactly ($r = 1.0$) to those obtained from the first dimension of the simple CA of the affiliation matrix for both the row and column objects.

@@

\begin{figure}[ht!]
    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
    \centering
     \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/ca-reord.png}
            \caption{Re-ordered Affiliation matrix (CA First Dimension).}
            \label{fig:ca-reord}
    \end{subfigure}
     \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/bon-reord1.png}
            \caption{Re-ordered Affiliation matrix (Bonacich Centrality).}
            \label{fig:bon-reord`}
    \end{subfigure}
     \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/bon-reord2.png}
            \caption{Re-ordered Affiliation matrix (Bonacich Second Dimension).}
            \label{fig:bon-reord}
    \end{subfigure}
    \caption{Southern Women Affiliation matrix with re-ordered rows and columns: (a) row and columns re-ordered according to the person and group score on the first CA dimension, (b) row and columns re-ordered according to person and group Eigenvector centrality score, (c) row and columns re-ordered according to person and group score on the second Bonacich dimension.}
    \label{fig:ca-v-bon}
\end{figure}

\subsection{Row and Column Re-ordered Affiliation Matrices} \label{sec:affmats}
Figures~\ref{fig:ca-v-bon}(a) and~\ref{fig:ca-v-bon}(b) illustrate the key differences between the two versions of dual centrality $C^R$ and $C^B$. Each panel shows the original Southern Women affiliation matrix but with rows and columns re-ordered according to the first CA dimension in (a) and the \citet{bonacich1991simultaneous} centrality in (b). In the plot, a cell entry is colored red if it has a one in the original affiliation matrix and is white when the corresponding entry is zero. As we can see, the two row-column-reshuffled matrices have appreciably distinct structures, with the $C^R$ re-ordered affiliation matrix having a block-diagonal structure and the $C^B$ re-ordered affiliation matrix having a triangular structure. Accordingly, the $C^R$ reflective scores reveal a dual community partition between groups of women who selectively attend two groups of events (on the top-right and bottom-left of the plot).

The traditional eigenvector centrality re-ordering, on the other hand, reveals a classic core-periphery partitioning \citep{borgatti2000models}, with a group of highly active women (on the top-left) who selectively co-participate in highly attended events (on the top right-hand side of the plot) and less active women (in the bottom half) who go to less well-attended events (on the left side). In fact, as \citet[p. 206-207]{everett2013dual} have noted, computing the eigenvector scores for both sets of nodes in the mode network from the classic Breiger projections---as in \citet{bonacich1991simultaneous}---and using them to re-order the rows and columns of the original affiliation matrix is a way to approximate using continuous scores an ideal discrete core-periphery partitioning of the two-mode data matrix, cheaper and more efficiently compared to direct optimization methods.  The larger point is that the standard eigenvector and reflective scores extract qualitatively different information from the two-mode network, with $C^R$ geared toward community partitioning and $C^B$ more focused on finding a core of well-connected people and groups.

\subsection{Eigenvector Plot} \label{subsec:eigplot}
We can confirm that the first CA axis points toward the underlying community partitioning of the two-mode network by looking at Figure~\ref{fig:ca-eigvec}(a) and~\ref{fig:ca-eigvec}(c), which shows the $C^R$ score of persons and groups on the x-axis against the score's rank order on the y-axis. If a two-mode network has no discernible community structure, the first dimension CA scores would be distributed as a continuous logistic curve; when community structure is present, we should observe discernible breaks in this distribution \citep{van2021correspondence}. The Southern Women data clearly belong to the second category. In the people mode, we have $\left[Frances, Laura, Brenda, Charlotte, Evenly, Theresa, Elanor, Ruth, Pearl\right]$ on one side, $\left[Dorothy, Nora, Katherine, Sylvia, Flora, Olivia, Myra, Helen, Verne\right]$ on the other. Among events, we have $\left[E1, E2, E3, E4, E5, E6\right]$ on side $\left[E9, E10, E11, E12, E13, E14\right]$ on the other and $\left[E7, E8\right]$ in an ambiguous middle position. Note that this is roughly the same partitioning obtained by \citep{doreian2004generalized} and reproduced by \citet{kovacs2010generalized} and \citet{lizardo2024two} using a generalized relational similarity approach. This indicates that persons and groups receive similar scores in the first dimension of CA only when they have similar connectivity similar to \textit{similar} groups, where group similarity is defined dually as having members in common who are themselves similar.

\begin{figure}
    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
    \centering
     \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/p-ca-eigvec.png}
            \caption{Eigenvector plot (Persons).}
            \label{fig:ca-reord}
    \end{subfigure}
     \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/g-ca-eigvec.png}
            \caption{Eigenvector plot (Groups).}
            \label{fig:bon-reord}
    \end{subfigure}
    \caption{CA Eigenvector plot for persons (a) and groups (b). The first CA score is on the x-axis, and the rank order of persons and groups is on the y-axis. Insets show the eigenvalue orderings on the x-axis ($k$) for persons and groups.}
    \label{fig:ca-eigvec}
\end{figure}

>>>>

\section{CA and Dual Scoring in Two-Mode Networks} \label{sec:ca}
The reader may ask what the point of going through all these reflective scoring since \citet{bonacich1991simultaneous} already developed a dual conception of scoring nodes in two-mode networks based on a very similar idea: Define the scores of entities in one mode based on the scores of entities in the other mode to which they are connected. Recall that, as noted earlier in that paper, Bonacich also considered CA in passing as a possible technique to assign scores to nodes in a two-mode network but (perhaps rightly) dismissed it as producing scores that could be interpreted as a kind of ``centrality,'' noting that it was better at capturing the clique structure of women who attended different events \citeyearpar[164]{bonacich1991simultaneous}.

While using CA to capture patterns of similarity in attendance between persons and groups makes sense (although the exact patterns of similarity captured by CA are not reducible to ``cliques'' as traditionally understood), CA can also be used to assign meaningful scores to nodes in a social network, which in some applications may even be interpreted as a type of centrality (or at least have a specifiable mathematical relation to scores that are usually unambiguously interpreted as centrality measures like degree). In fact, in the two-mode case, the first CA dimension captures the limit $(q \rightarrow \infty)$ of the reflective centralities discussed in the previous section \citep{mealy2019interpreting}. This makes the connection between the Bonacich and the CA scoring approaches more intimate than Bonacich thought.

To begin to see the linkages between CA and the Bonacich approach, recall that the solutions to the following system of linear equations give the Bonacich $(C^B)$ two-mode centralities:

@@

Where $A$ is the original affiliation matrix. Equations~\ref{eq:bon1} and~\ref{eq:bon2} have the typical form of a linear eigensystem, which means the unknown $C^B_g$ and $C^B_p$ scores can be obtained from the row and column eigenvectors corresponding to the largest eigenvalue $\lambda$, after performing a singular value decomposition (SVD) of the rectangular affiliation matrix $A$. Importantly, as Faust \citeyearpar[170]{faust1997centrality} notes, it is possible to express Equations~\ref{eq:bon1} and~\ref{eq:bon2} in terms of person and group-specific scores.

@@

Where $\lambda$ is the (first) eigenvalue corresponding to the eigenvector containing the $C^B_p$ scores. These equations make clear that in the Bonacich dual-scoring approach, the scores assigned to people are equal to the sum of the scores of the groups they join, and the scores assigned to groups are equivalent to the sum of the scores simultaneously assigned to their members. This approach thus nicely captures the duality property since the scores of nodes in each mode are given by aggregating their connections to nodes in the other mode.

However, Note the formal similarity between equations~\ref{eq:faust1} and~\ref{eq:faust2} and equations~\ref{eq:Rq_p} and~\ref{eq:Rq_g}. The main difference is that in HH's reflective equations, a person's centrality is proportional to the \textit{group-degree-weighted} centralities of the groups they join, and a group's centrality is proportional to the \textit{person-degree-weighted} centrality of the people who are members. We will return to this crucial point later.

\subsection{The Bonacich Two-Mode Centrality as a Reflective Centrality} \label{subsec:bonref}
To see the connection between the CA and Bonacich dual scores more clearly, we can motivate the Bonacich eigenvector-based approach using the same ``reflective'' exposition we used to introduce the HH reflective centrality in Section~\ref{sec:ref2mode}. Admittedly, this is a somewhat unorthodox way of presenting the eigenvector centralities for two-mode networks; however, it will help clarify the similarities and differences between the HH and the Bonacich approaches.

Accordingly, starting with Equations~\ref{eq:R1_p} and~\ref{eq:R1_g}, we can define a second-order ``Bonacich-reflection'' on both the persons and groups using the formulas:

@@

Where the denominator in~\ref{eq:BRq_pn} and~\ref{eq:BRq_gn} is the Euclidean vector norm.\footnote{For any vector $\mathbf{x}$ of length $n$, the $L_2$ norm is given by: $||\mathbf{x}||_2 = \sqrt{\sum_i^n x_i^2}$.} The normalization will prevent divergence of the sum of centralities for persons and groups, formalizing the weaker assumption of \textit{proportionality} between the centrality of each set of nodes and the sum of the centralities of the nodes in the other mode to which they are connected rather than the stronger assumption of strict equivalence \citep{bonacich_lloyd01}. Furthermore, the normalization guarantees convergence and the ``freezing'' of the sums of sums around steady values after a few iterations. These values will be equivalent (up to rounding error) to the (absolute value) of the dominant row and column eigenvectors of $\mathbf{A}$ as given in~\ref{eq:bon1} and~\ref{eq:bon2}.\footnote{Note that dividing by any vector norm---e.g., the $L_1$ or max norm---will prevent divergence and return scores perfectly correlated to the Bonacich eigenvector approach. Dividing by the $L_2$ norm returns scores that are \textit{exactly} the same, save for rounding error, as the absolute value of the dominant left and right eigenvectors of the affiliation matrix.} In fact, the iterative (normalized sum of sums) approach is one way of computing the leading eigenvectors of a rectangular matrix (e.g., the ``power method'' of \citet{mises1929praktische}).

This exercise establishes that there is more than a superficial similarity between the HH ``reflective'' scores and the Bonacich two-mode eigenvector centralities. In fact, both can be seen as instantiating an underlying ``prismatic'' model---in the sense of \citet{podolny2001networks}---of how centrality is distributed in the two-mode network. The Bonacich approach, based on sums of sums, favors overall \textit{volume}---or two-mode ``total effects'' in the sense of \citet{friedkin1991theoretical}---thus, groups receive more centrality points from their more central members and fewer centrality points from their less central members. Likewise, people receive more centrality points from their more central memberships and fewer centrality points from their less central memberships, and so on through every iteration. The HH reflective approach (and CA by implication) favors \textit{averages} rather than volume; thus, a person with a few memberships can contribute as much centrality to the group they belong to as a person with many memberships if the person with fewer memberships belongs to (on average) more central groups and the person with more memberships belongs to (on average) less central ones. In the same way, a person can receive as much centrality from a small group they belong to compared to a large group, as long as the small group (on average) has more central members in it, and the large group (on average) has less central members.

\subsection{HH Reflections as Correspondence Analysis} \label{subsec:refeigen}
Since iterating through normalized sums of sums is one way of obtaining the Bonacich two-mode centralities, it would be surprising if the equilibrium values of the HH reflective iterations were not themselves the solution to some eigenvalue decomposition problem. As has been noted recently by \citet{mealy2019interpreting} and \citet{van2021correspondence}, the Hidalgo and Hausmann's \citeyearpar{hidalgo2009building} reflective centralities can indeed be obtained directly (without iterations) as the solution to an eigenvalue decomposition problem.

@@

\begin{equation}
    \left(\mathbf{AA}^T\right)C^B_p = \lambda C^B_p
    \label{eq:bon3}
\end{equation}

\begin{equation}
    \left(\mathbf{A}^T\mathbf{A}\right)C^B_g = \lambda C^B_g
    \label{eq:bon4}
\end{equation}

This shows that the dual centrality Bonacich scores for persons and groups are equivalent to the eigenvectors of the respective one-mode projection matrices corresponding to the first (largest) eigenvalue, which qualifies the Bonacich scores as a ``dual projection'' method for analyzing two-mode networks in the sense of \citet{everett2013dual}.

We can motivate CA using the same dual projection strategy. Typically, when CA is presented as the solution to an eigenvalue problem, this is done by showing how the relevant scores can be obtained via the singular value decomposition (SVD) of the rectangular affiliation matrix~\cite{bonacich1991simultaneous, faust2005using}, which gives the impression that CA is exclusively a ``direct'' rather than a dual projection method. Here, we will show that CA can also be motivated, and scores obtained via a similar ``dual'' projection approach just like the Bonacich eigenvector scores, except that these are scores associated with a \textit{degree-weighted} projection of the original affiliation matrix.

To see this, consider the $|P| \times |P|$ matrix $\mathbf{D}p$, containing the ``first order'' reflective centralities of each person $C^R(1)_p$ (activity) along the diagonals and zeroes in every other cell. In the same way, consider the $|G| \times |G|$ matrix $\mathbf{D}g$, which contains the ``first order'' (degree) centralities of each group $C^R(1)_g$ (size) along the diagonals and zeroes in every other cell. Using these matrices, we can compute \textit{row stochastic} versions of the affiliation matrix and its transpose (recall that a matrix is row stochastic if its rows sum to one).

@@

We can do the same with the groups, except that we pre-multiply the \textit{transpose} of the original affiliation matrix by $\mathbf{D}g^{-1}$ (which now contains the \textit{inverse} of the first-order centralities of each group $C^R_g(1)$ along the diagonals) yielding the row-stochastic matrix $\mathbf{P}_{gp}$ of dimensions $|G| \times |P|$:

@@

Both $\mathbf{P}_{pp}$ and $\mathbf{P}_{gg}$ are square ($|P| \times |P|$ and $|G| \times |G|$) respectively) degree-normalized projection matrices linking people and groups. Like their constituent matrices, both $\mathbf{P}_{pp}$ and $\mathbf{P}_{gg}$ are row-stochastic (rows constrained to sum to 1.0), positive definite, but not symmetric, as can be seen in Figure~\ref{fig:dwproj}. Interestingly, one way to interpret the entries in the degree-weighted projection matrices for people and groups is by giving the \textit{probabilities} that a random walker starting at the row node and, following any sequence of $person-group-person'-group'$ hops, will reach the column person. Thus, higher values indicate an affinity or proximity between the people based on indirect connections in the two-mode network (and the same for groups in the corresponding matrix). The reason for the asymmetry is that depending on their pattern of connections to other groups, person $i$ may have more ways of indirectly reaching person $j$ than vice versa (unless $i$ and $j$ are connected to the same groups).

@@

With the $C^R$ scores for persons and groups obtained from the eigenvectors corresponding to the \textit{second} largest eigenvalue of the respective degree-weighted projection matrices.\footnote{Note also that the degree normalized projection matrices have equivalent sets of (non-zero) eigenvalues equal to the rank of the original affiliation matrix, with values between zero and one \citep{van2021correspondence}.}  This is because, as we have seen, both equations~\ref{eq:ca3} and~\ref{eq:ca4} produce row stochastic matrices; this means that their largest (dominant) eigenvalue will also be 1.0, which will be associated with a trivial first eigenvector containing constant values for both the people and the groups.  Accordingly, the scores we seek, which are the same as the CA scores and the limit HH reflective centrality scores, are those stored in the eigenvectors associated with the second largest eigenvalue \citep{van2021correspondence}; these will be equivalent to the row and column scores corresponding to the first non-trivial dimension (for people and groups respectively) obtained from a simple CA of the original affiliation matrix \citep[398, eq. 9.17]{fouss2016algorithms} or via the eigenvalue decomposition of the weighted projection matrices as in~\ref{eq:dam1} and~\ref{eq:dam2}. And as we can see in Figure~\ref{fig:ca-ref-corr}, the scores computed via the iterative method of reflections ($q = 26$) correspond exactly ($r = 1.0$) to those obtained from the first dimension of the simple CA of the affiliation matrix.

More importantly, because $\mathbf{P}_{pp} = \mathbf{D}p^{-1}\mathbf{A}\mathbf{D}g^{-1}\mathbf{A}^T$ and $\mathbf{P}_{gg} = \mathbf{D}g^{-1}\mathbf{A}^T\mathbf{D}p^{-1}\mathbf{A}$ there is a formal equivalence in the duality relations implied by equations ~\ref{eq:dam1} and~\ref{eq:dam2} and that implied by the Bonacich two-mode centralities in equations~\ref{eq:bon3} and~\ref{eq:bon4}. Both extract individual and group centralities as eigenvectors of the one-mode projection of the original affiliation matrix: $\mathbf{AA}^T$ and $\mathbf{P}_{pp}$ for people and $\mathbf{A}^T\mathbf{A}$ and $\mathbf{P}_{gg}$ for groups. The key difference is that the reflective centralities pre-multiply the affiliation matrix and its transpose by the inverse of the first-order centralities of the nodes in each mode before computing the eigenvalue decomposition, essentially normalizing the one-mode projection by the degrees of both sets of nodes.

This can be clearly seen if we express $\mathbf{P}_{PP}= \mathbf{D}g^{-1}\mathbf{A}^T\mathbf{D}p^{-1}\mathbf{A}$ in terms of each cell entry \cite[eq. 4]{mealy2019interpreting}:

@@

The same reasoning applies to groups in the matrix $\mathbf{P}_{gg} = \mathbf{D}g^{-1}\mathbf{A}^T\mathbf{D}p^{-1}\mathbf{A}$, whose entries are given by:

@@

Now, as has been noted by two-mode data analysts in other contexts \citep[e.g.,][]{faust2005using} it turns out that this ``double-pre-weighting'' of each cell entry by the degrees of each mode (e.g., the row and column sums of the original affiliation matrix) is precisely that used for the CA of a two-mode matrix.\footnote{More accurately, cells are weighted by the inverse of the square root of the product of the row, and column sums \citep[e.g.,][124]{faust2005using}; accordingly, the first non-trivial CA dimension for persons and groups given by the right and left eigenvectors corresponding to the second-largest eigenvalue from the SVD of the matrix $\mathbf{D}_p^{-1/2}\mathbf{A}\mathbf{D}_g^{-1/2}$ \citep[126]{faust2005using}.} In fact, the method of ``reflections'' is just a re-discovery of the older idea of reciprocal averaging \citep{hill1973reciprocal}, which is yet another way of computing row and column scores for elements in a two-mode matrix that are substantively equivalent to those obtained by (the first non-trivial dimension of) CA \citep{mealy2019interpreting}.

@@

\begin{figure}[ht!]
    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
    \centering
     \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/ca-reord.png}
            \caption{Re-ordered Affiliation matrix (CA First Dimension).}
            \label{fig:ca-reord}
    \end{subfigure}
     \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/bon-reord1.png}
            \caption{Re-ordered Affiliation matrix (Bonacich Centrality).}
            \label{fig:bon-reord1}
    \end{subfigure}
     \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/bon-reord2.png}
            \caption{Re-ordered Affiliation matrix (Bonacich Second Dimension).}
            \label{fig:bon-reord2}
    \end{subfigure}
    \caption{Southern Women Affiliation matrix with re-ordered rows and columns: (a) row and columns re-ordered according to the person and group score on the first CA dimension, (b) row and columns re-ordered according to person and group Eigenvector centrality score, (c) row and columns re-ordered according to person and group score on the second Bonacich dimension.}
    \label{fig:ca-v-bon}
\end{figure}

\subsection{Comparing Bonacich and CA Dual Scoring in the Southern Women Data}
\subsubsection{Row and Column Re-ordered Affiliation Matrices} \label{sec:affmats}
Figures~\ref{fig:ca-reord} and~\ref{fig:bon-reord1} illustrate the key differences between the two versions of dual centrality $C^R$ and $C^B$. Each panel shows the original Southern Women affiliation matrix but with rows and columns re-ordered according to the first CA dimension in (a) and the \citet{bonacich1991simultaneous} centrality in (b). In the plot, a cell entry is colored red if it has a one in the original affiliation matrix and is white when the corresponding entry is zero. As we can see, the two row-column-reshuffled matrices have appreciably distinct structures, with the $C^R$ re-ordered affiliation matrix having a block-diagonal structure and the $C^B$ re-ordered affiliation matrix having a triangular structure. Accordingly, the $C^R$ reflective scores reveal a dual community partition between groups of women who selectively attend two groups of events (on the top-right and bottom-left of the plot).

The traditional eigenvector centrality re-ordering, on the other hand, reveals a classic core-periphery partitioning \citep{borgatti2000models}, with a group of highly active women (on the top-left) who selectively co-participate in highly attended events (on the top right-hand side of the plot) and less active women (in the bottom half) who go to less well-attended events (on the left side). In fact, as \citet[p. 206-207]{everett2013dual} have noted, computing the eigenvector scores for both sets of nodes in the mode network from the classic Breiger projections---as in \citet{bonacich1991simultaneous}---and using them to re-order the rows and columns of the original affiliation matrix is a way to approximate using continuous scores an ideal discrete core-periphery partitioning of the two-mode data matrix, cheaper and more efficiently compared to discrete optimization methods. The larger point is that the standard eigenvector and reflective scores extract qualitatively different information from the two-mode network, with $C^R$ geared toward community partitioning and $C^B$ more focused on finding a core of well-connected people and groups.

\subsubsection{Eigenvector Plot} \label{subsec:eigplot}
We can confirm that the first CA axis points toward the underlying community partitioning of the two-mode network by looking at Figure~\ref{fig:ca-eigvec}(a) and~\ref{fig:ca-eigvec}(c), which shows the $C^R$ score of persons and groups on the x-axis against the score's rank order on the y-axis. If a two-mode network has no discernible community structure, the first dimension CA scores would be distributed as a continuous logistic curve; when community structure is present, we should observe discernible breaks in this distribution \citep{van2021correspondence}. The Southern Women data clearly belong to the second category. In the people mode, we have $\left[Frances, Laura, Brenda, Charlotte, Evenly, Theresa, Elanor, Ruth, Pearl\right]$ on one side, $\left[Dorothy, Nora, Katherine, Sylvia, Flora, Olivia, Myra, Helen, Verne\right]$ on the other. Among events, we have $\left[E1, E2, E3, E4, E5, E6\right]$ on side $\left[E9, E10, E11, E12, E13, E14\right]$ on the other and $\left[E7, E8\right]$ in an ambiguous middle position. Note that this is the canonical consensus partitioning (for the people) highlighted by \citet{freeman2003finding} across twenty-one different methods and subsequently reproduced by \citet{kovacs2010generalized} and \citet{lizardo2024two} using a generalized relational similarity (GRS) approach. This indicates that persons and groups receive similar scores in the first dimension of CA only when they have similar connectivity similar to \textit{similar} groups, where group similarity is defined dually as having members in common who are themselves similar.

\begin{figure}
    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
    \centering
     \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/p-ca-eigvec.png}
            \caption{Eigenvector plot (Persons).}
            \label{fig:p-eigvec}
    \end{subfigure}
     \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/g-ca-eigvec.png}
            \caption{Eigenvector plot (Groups).}
            \label{fig:g-eigvec}
    \end{subfigure}
    \caption{CA Eigenvector plot for persons (a) and groups (b). The first CA score is on the x-axis, and the rank order of persons and groups is on the y-axis. Insets show the eigenvalue orderings on the x-axis ($k$) for persons and groups.}
    \label{fig:ca-eigvec}
\end{figure}
